\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants determine interactions with floral visitors

### 2022-05-21 Manuscript/Lanuza_et_al_2021.Rmd
edits intro

header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{setspace}
- \usepackage{titlesec}
- \titlespacing{\title}{0pt}{\parskip}{-\parskip}
- \usepackage[style=nature]{biblatex}
- \usepackage[labelformat = empty]{caption}
- \usepackage{pdflscape}
- \usepackage[T1]{fontenc}
- \DeclareUnicodeCharacter{0301}{*************************************}
- \DeclareUnicodeCharacter{1EF3}{*************************************}

---

@@

\textbf{Jose B. Lanuza$^{1,2}$} <barragansljose@gmail.com>, \textbf{Romina Rader$^{1}$} <rrader@une.edu.au>, \textbf{Jamie Stavert$^{3}$} <jamie.stavert@gmail.com>, \textbf{Liam K. Kendall$^{4}$} <liam.k.kendall@gmail.com>, \textbf{Manu E. Saunders$^{1}$} <Manu.Saunders@une.edu.au> \textbf{and Ignasi Bartomeus$^{2}$} <nacho.bartomeus@gmail.com>

@@

^1^ School of Environmental and Rural Science, University of New England, Armidale, New South Wales 2350, Australia. ^2^ Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain. ^3^ Department of Conservation | Te Papa Atawhai, Auckland, New Zealand. ^4^ Centre for Environmental and Climate Science, Lund University, Sölvegatan 37, S-223 62 Lund, Sweden.

\doublespacing
\normalsize

\textbf{Statement of authorship:} JBL, RR and IB designed the study. JBL collated the data and conducted analysis with guidance of JS, LKK and IB. JBL wrote the manuscript with contributions of all authors.

@@

\textbf{Keywords:} life-history strategies | plant reproductive trade-offs | plant-pollinator interactions.

@@

\textbf{Corresponding author:} Jose B. Lanuza | +34 616375981| <barragansljose@gmail.com>

@@

\doublespacing
\vspace{5mm}
\normalsize

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

@@

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape plant-pollinator associations [@fenster2004; @dellinger2020]. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [@salguero2016]. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits [@friedman2020; @roddy2021], their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits [@wright2004; @chave2009; @diaz2016; @onoda2017; @laughlin2021], there is a need to acquire a multitrait perspective with broad geographical coverage that allows elucidating the reproductive spectrum of trait variation of flowering plants.

Although there is an increasing number of macroecological studies that investigate plant reproductive traits [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017], we still have poor understanding of how reproductive traits drive interactions with floral visitors at large ecological scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. In addition, the pollination system of a great number of plant species remains unexplored and is still unclear how specific key reproductive traits like mating or compatibility system influence plant-pollinator associations [@tur2013; @devaux2014]. Interestingly, the use of trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016] has shown to be of great importance when exploring the drivers of plant-pollinator interactions. Because the species' morphology can determine the species' functional role, an interesting novel approach to investigate plant-pollinator interactions is to examine how traits in the multidimensional trait space determine species interaction patterns [see @dehling2016]. Thus, by defining the reproductive spectrum of trait variation is possible to investigate how the different reproductive strategies are associated worldwide with the interaction patterns of the different floral visitors.

[Keep from here](Jose)

With the recent availability of large trait databases [e.g., TRY @kattge2011; and COMPADRE @salguero2015], plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function [@salguero2016; @diaz2016; @bruelheide2018; @carmona2021]. However, most studies have focused on vegetative traits such as leaf [@wright2004], wood [@chave2009], or root [@laughlin2021] trade-offs with little or no attention given to reproductive traits [@evojtko2020; @roddy2021] which are critical to plant life strategies that shape interactions with floral visitors and ultimately determine plant reproductive success.

For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively, [@barrett2003; @moeller2017] and outcrossing levels are positively correlated with flower size [@goodwillie2010]. In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species’ reliance upon animal pollination (pollinator dependence) and its role in attracting pollinators [@ollerton2011; @rodger2021]. However, it is still unknown to what extent these different reproductive strategies determine plant-pollinator interactions.

Several studies have identified links between plant traits and plant-pollinator network properties [@bartomeus2013; @olito2015; @rowe2020]. Moreover, plant traits can define species’ network roles [e.g., specialists vs generalists\; @lazaro2013; @tur2013]. For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators [@junker2013; @coux2016]. Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure [@stang2009; @ibanez2012]. Importantly, the combination of traits has shown to increase the predictive power of the network interactions [@eklof2013] and by considering the position in the multidimensional trait space we can determine species’ roles within plant-pollinator interaction networks [@dehling2016]. Nonetheless, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.

Here, we aim to explore the potential trade-offs among plant reproductive traits and their association with the different floral visitor guilds. First, we identify the major axes of reproductive trait variation and trade-offs that determine plant form and function. Second, we investigate the association between the plant species’ position in trait-space and the different floral visitor guilds with the help of qualitative (presence-absence of interaction) and quantitative (number of visits) information. Finally, we investigate how both the main axes of trait variation, and individual traits, influence plant species’ roles within networks using a set of complementary interaction network metrics (i.e., number of visits, normalized degree and specialization).

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks (see Table S1 and Fig. S1). These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies [e.g.,\ @carvalheiro2014; @olesen2007; @fortuna2010] or available in online archives [e.g., The Web of Life, @fortuna2010; and Mangal, @poisot2016]. In total, our network dataset constituted 60 weighted (number of visits) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (https://www.ncbi.nlm.nih.gov/taxonomy) for plants and ITIS (https://www.itis.gov/) for pollinators, using the R package _taxize_ [@chamberlain2020]. We filled the ‘not found’ searches manually using http://www.theplantlist.org/ and http://www.mobot.org/ for plants and http://www.catalogueoflife.org/ for floral visitors.

@@

We calculated the phylogenetic distance between different plant species using the function *get_tree* from the package _rtrees_ (https://github.com/daijiang/rtrees), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life [@smith2018; @jin2019].

@@

Trait missing values were imputed with the function _missForest_ [@stekhoven2012] which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of _missForest_ [@penone2014]. To extract the eigenvectors, we used the function _PVRdecomp_ from the package _PVR_ [@santos2018] based on a previous conceptual framework that considers phylogenetic eigenvectors [@diniz-filho2012]. We conducted two different imputations, one for the full set of species (1,506 species, 5.79% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (Fig. S2) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01% of missing values).

@@

We explored the trade-offs between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson’s correlation = 0.72, _P_ < 0.01 and Pearson’s correlation = 0.36, _P_ < 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the association of the different qualitative traits with the main axes of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers, we excluded values within the 2.5th–97.5th percentile range [@legendre2012], and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function _phyl.pca_ from the package _phytools_ [@revell2012] with the method lambda ($\lambda$) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation [@abdi2010]. Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S3). We found little difference between the explained variance with the imputed dataset (51.08%) and the dataset without missing values (52.87%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

@@

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package _phytools_ [@revell2012] and we used Pagel’s $\lambda$ as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

We first investigated how the different groups of floral visitors interacted along the main axes of trait variation (see below ‘visitation patterns’ section). This was done for the binary version of the networks that assumes equal weight across interactions and for the weighted version with the number of visits of floral visitors to individual flowers that account for the intensity of the interaction. Although floral visitors are not always pollinators and number of visits does not consider each pollinator species efficiency [@ballantyne2015], the number of visits can provide valuable information of the contribution of floral visitors to pollination [@vazquez2005; @vazquez2012]. Second, we investigated how the main axes of trait variation and individual traits influence plant species’ roles within networks using a set of complementary interaction network metrics: number of visits, normalized degree and specialization (see below ‘plant species network roles’ section).

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version _4.0.3_.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and number of visits of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that _Apis mellifera_ was the floral visitor with the largest proportion of records counted (7.55% of the total). This finding is consistent with previous research showing that _A. mellifera_ was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of _A. mellifera_ on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding _A. mellifera_. We found that _A. mellifera_, was partly driving some of the observed trends on PC1 (Fig. S4). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package _brms_ [@burkner2017]. We modelled presence-absence of observed interactions and number of visits as a function of the main axes of plant trait variation and their interactions with floral visitor guilds (e.g., number of visits ~ PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different floral visitor guilds. We added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified for presence-absence of interaction and number of visits a Bernoulli and a zero inflated negative binomial distribution, respectively. The models were run with non or very weakly informative informative priors from the brm function so they have neglible influence on the results [@burkner2017], 3,000 iterations and with previous 1,000 warm up iterations. We set delta ($\Delta$) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function *pp_check* using the _bayesplot_ package [@gabry2019].

@@

We investigated whether different quantitative traits determined plant species network roles using Bayesian modelling and regression trees. For this, we selected simple and complementary species-level network metrics commonly applied in bipartite network studies [@dormann2008] with a straightforward ecological interpretation relevant to our research goals. The different plant species-level metrics were: (i) sum of visits per plant species; (ii) normalized degree, calculated as the number of links per plant species divided by the total possible number of partners; and (iii) specialization (d’) [@bluthgen2006], which measures the deviation of an expected random choice of the available interaction partners and ranges between 0 (maximum generalization) and 1 (maximum specialization). Normalized degree and specialization were calculated with the _specieslevel_ function from the R package _bipartite_ [@dormann2008].

First, we modelled the distinct plant species metrics (sum of visits, normalized degree and plant specialization) as a function of the three main axes of trait variation (plant species level metric ~ PC1 + PC2 + PC3). For each response variable (i.e., each plant species level metric), we used different distribution families (zero inflated negative binomial for the sum of visits, weibull for normalized degree and zero one inflated beta for specialization). Finally, we used the same random factors, model settings and conducted the same posterior predictive checks for each model as detailed above in the ‘visitation patterns’ section.

Second, to better understand complex trait relationships, we used regression trees. Regression trees are recursive algorithms which can detect complex relationships among predictors and allow identification of the relevance of specific trait combinations on explaining species roles within the network of interaction. We focused exclusively on quantitative traits because almost all categorical traits were statistically associated with the first two axes of trait variation (Table S2). We conducted this analysis using the _rpart_ function from the _rtrees_ package [@therneau2015] with method _‘anova’_ with a minimum of 50 observations per terminal node and we used the _rpart.plot_ package [@milborrow2015] to plot the regression trees. We considered the species level indices as response variables (number of visits, normalized degree and specialization) and we performed one regression tree per metric using the different quantitative traits as predictors. We calculated two regression trees per plant species-level metric, one for the full set of species and another for the subset of species for which we had pollen and nectar traits. We focused on regression trees that included floral rewards because they consistently showed pollen and nectar traits as being the best for explaining the different species-level metrics (Fig. S5).

@@

We found a strong phylogenetic signal (*P* < 0.01) in all quantitative traits (Table S5). The traits that showed the highest phylogenetic signal were ovule number ($\lambda$ = 1), pollen grains per flower ($\lambda$ = 1) and plant height ($\lambda$ = 0.96), followed by flower length ($\lambda$ = 0.75), flower width ($\lambda$ = 0.73), number of flowers per plant ($\lambda$ = 0.69) and nectar concentration ($\lambda$ = 0.65). The traits that showed a moderate phylogenetic signal were inflorescence width ($\lambda$ = 0.57), style length ($\lambda$ = 0.49) and autonomous selfing ($\lambda$ = 0.34). Finally, microliters of nectar per flower showed the lowest phylogenetic signal of all traits ($\lambda$ = 0.14).

@@

<div id ="refs"></div>

@@

\vspace*{-28mm}

>>>>

header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{setspace}
- \usepackage{titlesec}
- \titlespacing{\title}{0pt}{\parskip}{-\parskip}
- \usepackage[style=nature]{biblatex}
- \usepackage[labelformat = empty]{caption}
- \usepackage{pdflscape}
- \usepackage[T1]{fontenc}
- \DeclareUnicodeCharacter{0301}{*************************************}
- \DeclareUnicodeCharacter{1EF3}{*************************************}
---

@@

\textbf{Jose B. Lanuza$^{1,2}$} [barragansljose\@gmail.com](mailto:barragansljose@gmail.com){.email}, \textbf{Romina Rader$^{1}$} [rrader\@une.edu.au](mailto:rrader@une.edu.au){.email}, \textbf{Jamie Stavert$^{3}$} [jamie.stavert\@gmail.com](mailto:jamie.stavert@gmail.com){.email}, \textbf{Liam K. Kendall$^{4}$} [liam.k.kendall\@gmail.com](mailto:liam.k.kendall@gmail.com){.email}, \textbf{Manu E. Saunders$^{1}$} [Manu.Saunders\@une.edu.au](mailto:Manu.Saunders@une.edu.au){.email} \textbf{and Ignasi Bartomeus$^{2}$} [nacho.bartomeus\@gmail.com](mailto:nacho.bartomeus@gmail.com){.email}

@@

^1^ School of Environmental and Rural Science, University of New England, Armidale, New South Wales 2350, Australia. ^2^ Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain. ^3^ Department of Conservation \| Te Papa Atawhai, Auckland, New Zealand. ^4^ Centre for Environmental and Climate Science, Lund University, Sölvegatan 37, S-223 62 Lund, Sweden.

```{=tex}
\doublespacing
\normalsize
```
\textbf{Statement of authorship:} JBL, RR and IB designed the study. JBL collated the data and conducted analysis with guidance of JS, LKK and IB. JBL wrote the manuscript with contributions of all authors.

@@

\textbf{Keywords:} life-history strategies \| plant reproductive trade-offs \| plant-pollinator interactions.

@@

\textbf{Corresponding author:} Jose B. Lanuza \| +34 616375981\| [barragansljose\@gmail.com](mailto:barragansljose@gmail.com){.email}

@@

```{=tex}
\doublespacing
\vspace{5mm}
\normalsize
```
```{r setup, include=FALSE, cache=FALSE, message = FALSE}

@@

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape plant-pollinator associations [@fenster2004; @dellinger2020]. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [@stearns1989]. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits [@salguero2016; @friedman2020; @roddy2021], their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other vegetative and physiological plant traits [@wright2004; @chave2009; @diaz2016; @onoda2017; @laughlin2021], there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive trait covariation patterns can help to further understand the different plant ecological strategies [@agrawal2020] and improve our knowledge on plant-pollinator associations [@roddy2021].

Although there is an increasing number of macroecological studies that investigate plant reproductive traits [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017], we still have poor understanding of how reproductive traits drive interactions with floral visitors at large ecological scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. In addition, the pollination system of a great number of plant species remains unexplored and is still unclear how specific key reproductive traits like mating or compatibility system influence plant-pollinator associations [@tur2013; @devaux2014]. Interestingly, the use of trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016] has shown to be of great importance when exploring the drivers of plant-pollinator interactions. For example, plant traits can define species' network roles [e.g., specialists vs generalists; @lazaro2013; @tur2013] and plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators [@junker2013; @coux2016]. Indeed, morphological matching between plants and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure [@stang2009; @ibanez2012]. Because the species' morphology can determine the species' functional role in the pollination network and the combination of traits has shown to increase the predictive power of the network interactions [@eklof2013], an interesting novel approach is to investigate how traits in the multidimensional trait space determine species interaction patterns [see @dehling2016]. Thus, by exploring the reproductive spectrum of trait variation is possible to delimit the different plant reproductive strategies and explore how these are associated with the interaction patterns of the different floral visitors.

With the recent availability of large trait databases [e.g., TRY @kattge2011; and COMPADRE @salguero2015], plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function [@salguero2016; @diaz2016; @bruelheide2018; @carmona2021]. However, most studies with a multitrait perspective have focused on trait correlations from the leaf [@wright2004], wood [@chave2009], or root [@laughlin2021] related traits with little or no attention given to reproductive traits [@evojtko2020; @roddy2021]. Despite the lack of an holistic view that depict reproductive trait covariation patterns, there are widely recognized reproductive trait associations between pair of traits such as the negative correlation between flower size and flower number [@sargent2007; @kettle2011], the positive association between flower size and outcrossing rate [@goodwillie2010] or the association between outcrossing rate and lifespan where short lived versus perennial species tend to have low versus high levels of outcrossing [@barrett2003; @moeller2017], respectively. Although these different trait associations (and others) have recently allowed to progress towards a conceptual framework that integrates the different floral trait relationships [@roddy2021], we still lack empirical evidence that investigates jointly these different reproductive trait associations.

Here, we aim to progress knowledge on the reproductive trait covariation patterns and their association with the different floral visitor guilds by exploring at a broad geographical scale the reproductive trait associations of entomophilous plant species from plant-pollination networks. First, we investigate what are the major axes of reproductive trait variation for the different plant species. Second, we investigate the association between the plant species' position in the multidimensional trait-space and the different floral visitor guilds with the help of qualitative (presence-absence of interaction) and quantitative (number of visits) information about plant-pollinator interactions. Finally, we investigate how both the main axes of trait variation, and individual traits, influence plant species' functional role in the pollination network using a set of complementary interaction network metrics (i.e., number of visits, normalized degree and specialization).

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks (see Table S1 and Fig. S1). These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies [e.g.,  @carvalheiro2014; @olesen2007; @fortuna2010] or available in online archives [e.g., The Web of Life, @fortuna2010; and Mangal, @poisot2016]. In total, our network dataset constituted 60 weighted (number of visits) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (<https://www.ncbi.nlm.nih.gov/taxonomy>) for plants and ITIS (<https://www.itis.gov/>) for pollinators, using the R package *taxize* [@chamberlain2020]. We filled the 'not found' searches manually using <http://www.theplantlist.org/> and <http://www.mobot.org/> for plants and <http://www.catalogueoflife.org/> for floral visitors.

@@

We calculated the phylogenetic distance between different plant species using the function *get_tree* from the package *rtrees* (<https://github.com/daijiang/rtrees>), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life [@smith2018; @jin2019].

@@

Trait missing values were imputed with the function *missForest* [@stekhoven2012] which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of *missForest* [@penone2014]. To extract the eigenvectors, we used the function *PVRdecomp* from the package *PVR* [@santos2018] based on a previous conceptual framework that considers phylogenetic eigenvectors [@diniz-filho2012]. We conducted two different imputations, one for the full set of species (1,506 species, 5.79% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (Fig. S2) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01% of missing values).

@@

We explored the trade-offs between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, *P* \< 0.01 and Pearson's correlation = 0.36, *P* \< 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the association of the different qualitative traits with the main axes of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers, we excluded values within the 2.5th--97.5th percentile range [@legendre2012], and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function *phyl.pca* from the package *phytools* [@revell2012] with the method lambda ($\lambda$) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation [@abdi2010]. Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S3). We found little difference between the explained variance with the imputed dataset (51.08%) and the dataset without missing values (52.87%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

@@

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package *phytools* [@revell2012] and we used Pagel's $\lambda$ as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

We first investigated how the different groups of floral visitors interacted along the main axes of trait variation (see below 'visitation patterns' section). This was done for the binary version of the networks that assumes equal weight across interactions and for the weighted version with the number of visits of floral visitors to individual flowers that account for the intensity of the interaction. Although floral visitors are not always pollinators and number of visits does not consider each pollinator species efficiency [@ballantyne2015], the number of visits can provide valuable information of the contribution of floral visitors to pollination [@vazquez2005; @vazquez2012]. Second, we investigated how the main axes of trait variation and individual traits influence plant species' roles within networks using a set of complementary interaction network metrics: number of visits, normalized degree and specialization (see below 'plant species network roles' section).

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version *4.0.3*.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and number of visits of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that *Apis mellifera* was the floral visitor with the largest proportion of records counted (7.55% of the total). This finding is consistent with previous research showing that *A. mellifera* was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of *A. mellifera* on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding *A. mellifera*. We found that *A. mellifera*, was partly driving some of the observed trends on PC1 (Fig. S4). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package *brms* [@burkner2017]. We modelled presence-absence of observed interactions and number of visits as a function of the main axes of plant trait variation and their interactions with floral visitor guilds (e.g., number of visits \~ PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different floral visitor guilds. We added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified for presence-absence of interaction and number of visits a Bernoulli and a zero inflated negative binomial distribution, respectively. The models were run with non or very weakly informative informative priors from the brm function so they have neglible influence on the results [@burkner2017], 3,000 iterations and with previous 1,000 warm up iterations. We set delta ($\Delta$) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function *pp_check* using the *bayesplot* package [@gabry2019].

@@

We investigated whether different quantitative traits determined plant species network roles using Bayesian modelling and regression trees. For this, we selected simple and complementary species-level network metrics commonly applied in bipartite network studies [@dormann2008] with a straightforward ecological interpretation relevant to our research goals. The different plant species-level metrics were: (i) sum of visits per plant species; (ii) normalized degree, calculated as the number of links per plant species divided by the total possible number of partners; and (iii) specialization (d') [@bluthgen2006], which measures the deviation of an expected random choice of the available interaction partners and ranges between 0 (maximum generalization) and 1 (maximum specialization). Normalized degree and specialization were calculated with the *specieslevel* function from the R package *bipartite* [@dormann2008].

First, we modelled the distinct plant species metrics (sum of visits, normalized degree and plant specialization) as a function of the three main axes of trait variation (plant species level metric \~ PC1 + PC2 + PC3). For each response variable (i.e., each plant species level metric), we used different distribution families (zero inflated negative binomial for the sum of visits, weibull for normalized degree and zero one inflated beta for specialization). Finally, we used the same random factors, model settings and conducted the same posterior predictive checks for each model as detailed above in the 'visitation patterns' section.

Second, to better understand complex trait relationships, we used regression trees. Regression trees are recursive algorithms which can detect complex relationships among predictors and allow identification of the relevance of specific trait combinations on explaining species roles within the network of interaction. We focused exclusively on quantitative traits because almost all categorical traits were statistically associated with the first two axes of trait variation (Table S2). We conducted this analysis using the *rpart* function from the *rtrees* package [@therneau2015] with method *'anova'* with a minimum of 50 observations per terminal node and we used the *rpart.plot* package [@milborrow2015] to plot the regression trees. We considered the species level indices as response variables (number of visits, normalized degree and specialization) and we performed one regression tree per metric using the different quantitative traits as predictors. We calculated two regression trees per plant species-level metric, one for the full set of species and another for the subset of species for which we had pollen and nectar traits. We focused on regression trees that included floral rewards because they consistently showed pollen and nectar traits as being the best for explaining the different species-level metrics (Fig. S5).

@@

We found a strong phylogenetic signal (*P* \< 0.01) in all quantitative traits (Table S5). The traits that showed the highest phylogenetic signal were ovule number ($\lambda$ = 1), pollen grains per flower ($\lambda$ = 1) and plant height ($\lambda$ = 0.96), followed by flower length ($\lambda$ = 0.75), flower width ($\lambda$ = 0.73), number of flowers per plant ($\lambda$ = 0.69) and nectar concentration ($\lambda$ = 0.65). The traits that showed a moderate phylogenetic signal were inflorescence width ($\lambda$ = 0.57), style length ($\lambda$ = 0.49) and autonomous selfing ($\lambda$ = 0.34). Finally, microliters of nectar per flower showed the lowest phylogenetic signal of all traits ($\lambda$ = 0.14).

@@

::: {#refs}
:::

@@

\vspace{-28mm}
